---
title: "Should the Sixers Tank"
author: "Joe Yi"
date: "2025-02-07"
output: html_document
---

```{r libraries, include=FALSE}

library(dplyr)
library(rvest)
library(stringr)
```

## Should the Sixers Tank?
The point of this code is to figure out at what point are the sixers better off tanking or competing.

```{r webscrape, include=FALSE}

url <- "https://www.basketball-reference.com/leagues/NBA_2025_standings.html"

# Read the webpage’s HTML content
page <- read_html(url)

# Scrape the Eastern Conference standings table
east_table <- page %>% 
  html_node("table#divs_standings_E") %>% 
  html_table(fill = TRUE)

clean_east_table <- east_table %>% 
  filter(!grepl("Division", `Eastern Conference`)) %>% 
  mutate(Team = str_remove(`Eastern Conference`, "\\s*\\(\\d+\\)"))


# Scrape the Western Conference standings table
west_table <- page %>% 
  html_node("table#divs_standings_W") %>% 
  html_table(fill = TRUE)

clean_west_table <- west_table %>% 
  filter(!grepl("Division", `Western Conference`)) %>% 
  mutate(Team = str_remove(`Western Conference`, "\\s*\\(\\d+\\)"))

nba_standings <- bind_rows(clean_east_table, clean_west_table) %>% 
  select(Team, W, L) %>% 
  rename(
    Wins = W,
    Losses = L
  )

nba_standings <- nba_standings %>% 
  mutate(
    Wins = as.numeric(Wins),
    Losses = as.numeric(Losses)
  )

nba_standings$WLp <- nba_standings$Wins/(nba_standings$Wins + nba_standings$Losses)
```

## Setting up the Dataset

Ordering by Win/Loss percentage then using a randomizer to break ties.

Note we are using post-2019 NBA Draft Lottery Odds (worst 3 teams have equal odds)

```{r lottery, echo=FALSE}

nba_standings <- nba_standings %>% 
  arrange(WLp)

nba_standings_ties <- nba_standings %>% 
  group_by(WLp) %>% 
  mutate(tiebreaker = runif(n())) %>% 
  arrange(WLp, tiebreaker) %>% 
  ungroup()

lottery_teams <- nba_standings[1:14, ]

lottery_teams$lottery_odds <- c(0.14, 0.14, 0.14, 0.125, 0.105, 0.09, 0.075, 0.06, 0.045, 0.03, 0.02, 0.015, 0.01, 0.005)

```

```{r Function, echo=TRUE}

top_x_protected <- function(team_name, protection_limit, simulations = sims) {
  
  top_pick_count <- 0
  for (i in 1:simulations) {
    lottery_results <- sample(lottery_teams$Team, 4, prob = lottery_teams$lottery_odds, replace = FALSE)
    
    remaining_lottery_teams <- setdiff(lottery_teams$Team, lottery_results)
    
    draft_order <- c(lottery_results, remaining_lottery_teams, rev(nba_standings$Team[15:30]))
    
    pick_position <- which(draft_order == team_name)
    
    if (pick_position <= protection_limit) {
      top_pick_count <- top_pick_count +1
    
  
    }
  }
  
  protection_chance <- (top_pick_count / simulations) *100
  return(protection_chance)
}

```


```{r Chance, echo=TRUE}


sims = 1000000

chance <- top_x_protected("Philadelphia 76ers", 6)
print(chance)

# To see the sixer's chances if they were worse than the Nets, just replace the Team Name parameter with "Brooklyn Nets" instead of "Philadelphia 76ers"
```

## Conclusion

It's hopeless. Since the 76ers are at the 7th spot, they have to pray they get into the top 4.
